### 2024-03-11 docs/SOSP 2024/schedule.tex
Add xgb and Tahoe schedules

\section{Scheduling Language}
The goal of the scheduling language is to express the following
* The order in which the iteration space over the batch of inputs (batch) and trees in the forest (tree) is to be traversed. Note that there is a reduction over the tree dimension.
* Intra or inter tree optimizations that are to be performed on a tree or set of trees (tree walk unrolling, pipelining, SIMDize etc).

The reasons to use a scheduling language rather than a hard-coded lowering are as follows
* Making the scheduling specification external to the compiler allows us to 
more easily build auto-schedulers and auto-tuners.
* It is very hard to come up with a template loop nest that works for all models (for example, tree sizes may vary across the model making it necessary to iterate different number of trees at different times). 
* A scheduling language will make writing newer locality optimizations faster since no changes to the compiler infrastructure will be needed.
* Adding support for additional hardware targets (GPUs, FPGAs), will be much easier with a scheduling language.

There are some simplifying assumptions and limitations in the current design
* Tree traversals are considered atomic. There is no way to express partial tree traversals or schedule individual node/level computations. 
* Accumulation of tree predictions is done immediately (as opposed to, for example, collecting all predictions and performing a reduction later).

@@

>>>>

\section{Scheduling Language}
The goal of the scheduling language is to express the following
\begin{itemize}
  \item The order in which the iteration space over the batch of inputs (batch) and trees in the forest (tree) is to be traversed. Note that there is a reduction over the tree dimension.
  \item Intra or inter tree optimizations that are to be performed on a tree or set of trees (tree walk unrolling, pipelining, SIMDize etc).
\end{itemize}

The reasons to use a scheduling language rather than a hard-coded lowering are as follows
\begin{itemize}
  \item Making the scheduling specification external to the compiler allows us to 
  more easily build auto-schedulers and auto-tuners.
  \item It is very hard to come up with a template loop nest that works for all models 
  (for example, tree sizes may vary across the model making it necessary to iterate 
  different number of trees at different times).
  \item A scheduling language will make writing newer locality optimizations faster since 
  no changes to the compiler infrastructure will be needed.
  \item Adding support for additional hardware targets (GPUs, FPGAs), will be much easier
  with a scheduling language.
\end{itemize}

There are some simplifying assumptions and limitations in the current design
\begin{itemize}
  \item Tree traversals are considered atomic. There is no way to express partial
   tree traversals or schedule individual node/level computations. 
  \item Accumulation of tree predictions is done immediately (as opposed to, 
  for example, collecting all predictions and performing a reduction later).
\end{itemize}

@@

\subsection{The XBoost Schedule}
XGBoost\cite{XGBoost} is a very popular gradient boosting library. 
It implements inference on the CPU by going over a fixed number of rows (64 in the previous version)
for every tree and then moving to the next tree. When all trees have been walked for this 
set of rows, the next set of rows is taken up. Also, different sets of rows are 
processed in parallel.

The schedule used by XGBoost can be represented in \Treebeard{}'s scheduling 
language as follows.
\begin{lstlisting}[style=c++]
  tile(batch, b0, b1, CHUNK_SIZE)
  reorder(b0, tree, b1)
  parallel(b0)
\end{lstlisting}

\subsection{Tahoe Schedules}
Tahoe\cite{Tahoe} has four strategies that it picks from for a given model. 
Each of these strategies can be encoded using \Treebeard{}'s scheduling language
as we show below. 
\begin{itemize}
  \item \textbf{Direct Method}: In this strategy, a single GPU thread walks all trees
  for a given input row. The schedule for this strategy is as follows.
\begin{lstlisting}[style=c++]
  tile(batch, b0, b1, ROWS_PER_TB)
  reorder(b0, b1, tree)
  gpuDimension(b0, grid.x)
  gpuDimension(b1, block.x)
\end{lstlisting}
  Here, \op{ROWS\_PER\_TB} is the number of rows that are processed by a single thread block.
  \item \textbf{Shared Data}: In this strategy, a thread block walks all the trees 
  for a given row in parallel. If threads walk multiple trees, each thread accumulates
  partial results. Finally, a thread block wide reduction is performed to compute 
  the prediction. The schedule for this strategy is as follows.
\begin{lstlisting}[style=c++]
  reorder(batch, tree)
  gpuDimension(batch, grid.x)
  gpuDimension(tree, block.x)
  cache(batch)
\end{lstlisting}
  \item \textbf{Shared Forest}: In this strategy, the whole model is loaded into 
  shared memory and subsequently, a single thread walks all trees for a particular
  row. The schedule for this strategy is as follows.
\begin{lstlisting}[style=c++]
  tile(batch, b0, b1, ROWS_PER_TB)
  tile(tree, t0, t1, N_TREES)
  reorder(b0, b1, t0, t1)
  cache(t0)
  gpuDimension(b0, grid.x)
  gpuDimension(b1, block.x)
\end{lstlisting}
  Here, we create a placeholder single iteration loop \op{t0} so that we can 
  specify that all trees are to be cached.
  \item \textbf{Shared Partial Forest}: In case the model is too large to fit into
  shared memory, the model is split into chunks and each chunk is loaded into shared
  memory. Again, as in the previous strategy, one thread walks all trees assigned to
  a thread block for a row. The schedule for this strategy is as follows.
\begin{lstlisting}[style=c++]
  tile(batch, b0, b1, ROWS_PER_TB)

  tile(tree, t0, t0Inner, TREES_PER_TB)
  tile(t0Inner, t1, t2, TREES_PER_TB)
  cache(t1)
  reorder(b0, t0, b1, t1, t2)

  gpuDimension(b0, grid.x)
  gpuDimension(t0, grid.y)
  gpuDimension(b1, block.x)
\end{lstlisting}
\end{itemize}
